PANCHANGE_analysis

### 2020-08-19 PANCHANGE_analysis.Rmd
added calculated frequencies and plots

### Read in data
```{r read in data}
dat <- readRDS(file = paste0(data_path, "/four_cohorts.rds"))
dim(dat)
colnames(dat)
```

Filter data frame to exclude participants that never answered COPING
```{r}

```{r Vectors with diff_score_cols}
diff_score_cols <- dat %>%
  select(contains("diff_score")) %>% 
  colnames()

@@

```{r dat.coping summary, eval=FALSE, include=FALSE}
summarytools::dfSummary(
  dat,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```

Data availability
```{r gad availability}
dat %>%
  freq(
    data_group_gad,
    cumul = F)
```

```{r phq availability}
dat %>%
  freq(
    data_group_phq,
    cumul = F)
```

```{r pcl availability}
dat %>%
  freq(
    data_group_pcl,
    cumul = F)
```

Data availability OCI-R
```{r}

Descriptives of age
```{r}

```{r vector with psychiatric disorders}
disorder_columns <- c(
    "depressive_disorders_numeric",
    "anxiety_disorders_numeric",
    "eating_disorders_numeric",
    "obsessive_compulsive_disorders_numeric",
    "psychotic_disorders_numeric",
    "mhd.mania_hypomania_bipolar_or_manicdepression_numeric",
    "mhd.posttraumatic_stress_disorder_ptsd_numeric",
    "mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric",
    "mhd.attention_deficit_hyperactivity_disorder_numeric",
    "mhd.personality_disorder_numeric",
    "control_numeric"
)
```

```{r gather to long format Disorder}
dat.long <- dat %>%
  gather(
    key = "Disorder",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    all_of(diff_score_cols),
    Disorder,
    Diagnosis
  )

head(dat.long)
```

@@

```{r Disorder by Gender}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Gender) %>%
  count(Disorder)
```

```{r Disorder by Sample}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Sample) %>%
  count(Disorder)
```

Ethnicity should be recoded
European
Mixed or multiple ethnic origins
African or African British
Asian or Asian British
Arab
```{r Disorder by Ethnicity}
dat.long %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Ethnicity) %>%
  count(Disorder)
```

@@

Ethnicity
```{r}
dat %>%
  group_by(Ethnicity, Sample) %>%
  count(Ethnicity)
```

Gender
```{r}
dat %>%
  group_by(Gender, Sample) %>%
  count(Gender)
```

highest_education
```{r}
dat %>%
  group_by(highest_education, Sample) %>%
  count(highest_education)
```

Key worker status
```{r}
dat %>%
  group_by(key_worker, Sample) %>%
  count(key_worker)
```

COVID test positive
```{r}
dat %>%
  group_by(respiratory.throat_swab_test_nose, Sample) %>%
  count(respiratory.throat_swab_test_nose)
```

```{r}
dat %>%
  group_by(respiratory.results_antibody_test, Sample) %>%
  count(respiratory.results_antibody_test)
```

COVID probable diagnosis
```{r}
dat %>%
  group_by(covid_probable_case_incl_screening, Sample) %>%
  count(covid_probable_case_incl_screening)
```

```{r gather to long format Time_points}
dat.long.gad.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "gad.diff_score",
    all_of(gad.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Time_point_raw,
    gad.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "gad.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

makes this into percentages
```{r}
dat.long.gad.diff %>%
  ggplot(mapping = aes(y = gad.diff_score)) +
  geom_bar() +
  theme_minimal() +
  facet_wrap(facets = vars(Time_point))
```

```{r gather to long format Time_points}
dat.long.phq.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.diff_score",
    all_of(phq.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Time_point_raw,
    phq.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.diff_score_", "")) %>%
  select(-Time_point_raw) %>%
  drop_na(phq.diff_score)

>>>>

## Read in data
```{r read in data}
dat.raw <- readRDS(file = paste0(data_path, "/four_cohorts.rds"))
dim(dat.raw)
colnames(dat.raw)
```

```{r dat.coping summary, eval=FALSE, include=FALSE}
summarytools::dfSummary(
  dat.raw,
  graph.col = F,
  valid.col = F,
  labels.col = F)
```

# Data availability
```{r gad availability}
dat.raw %>%
  freq(
    data_group_gad,
    cumul = F)
```

```{r phq availability}
dat.raw %>%
  freq(
    data_group_phq,
    cumul = F)
```

```{r pcl availability}
dat.raw %>%
  freq(
    data_group_pcl,
    cumul = F)
```

Data availability OCI-R
```{r ocir availability}
dat.raw %>%
  freq(
    data_group_ocir,
    cumul = F)
```

# Exclusion
Filter data frame to exclude participants that never answered COPING
```{r exclusion criteria dat}
dat <- dat.raw %>%
  filter(
    (data_group_gad == "Baseline and retrospective data" |
      data_group_gad == "Prepandemic and baseline data" |
      data_group_gad == "All prepandemic, baseline and retrospective data") |
      (data_group_phq == "Baseline and retrospective data" |
      data_group_phq == "Prepandemic and baseline data" |
      data_group_phq == "All prepandemic, baseline and retrospective data") |
      (data_group_pcl == "Baseline and retrospective data" |
      data_group_pcl == "Prepandemic and baseline data" |
      data_group_pcl == "All prepandemic, baseline and retrospective data") |
      (data_group_ocir == "Baseline and retrospective data" |
      data_group_ocir == "Prepandemic and baseline data" |
      data_group_ocir == "All prepandemic, baseline and retrospective data")
  )
```

```{r dimensions of total dat}
dim(dat)
```

# ggplot theme
```{r ggplot theme}
theme_personal <-  theme(
    text = element_text(color = "black"),
    axis.title = element_text(color = "black"),
    axis.text = element_text(color = "black"),
    axis.title.y = element_blank(), 
    legend.background = element_blank(),
    legend.box.background = element_blank(),
    panel.background = element_blank(), 
    panel.grid.minor.x = element_blank(),
    panel.grid.major.x = element_line(
      colour = "gray",
      linetype = "dashed",
      size = 0.2
      ),
    axis.ticks = element_blank()
    )
```

# Age
```{r descriptives Ages}
dat %>%
  group_by(Sample) %>%
  summarise(Mean = mean(Age),
            SD = sd(Age),
            Min = min(Age),
            Max = max(Age))
```

# Vectors for diff_score
```{r Vectors with diff_score_cols}
diff_score_cols <- dat %>%
  select(contains("diff_score")) %>% 
  colnames()

@@

# Psychiatric disorders
```{r vector with psychiatric disorders}
disorder_columns <- c(
    "depressive_disorders_numeric",
    "anxiety_disorders_numeric",
    "eating_disorders_numeric",
    "obsessive_compulsive_disorders_numeric",
    "psychotic_disorders_numeric",
    "mhd.mania_hypomania_bipolar_or_manicdepression_numeric",
    "mhd.posttraumatic_stress_disorder_ptsd_numeric",
    "mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric",
    "mhd.attention_deficit_hyperactivity_disorder_numeric",
    "mhd.personality_disorder_numeric",
    "control_numeric"
)
```

```{r gather to long format Disorder}
dat.long.psych <- dat %>%
  gather(
    key = "Disorder_raw",
    value = "Diagnosis",
    all_of(disorder_columns)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    all_of(diff_score_cols),
    Disorder_raw,
    Diagnosis
  ) %>%
  mutate(Disorder =
           recode_factor(
             Disorder_raw,
             anxiety_disorders_numeric = "Anxiety disorders",
             depressive_disorders_numeric = "Depressive disorders",
             eating_disorders_numeric = "Eating disorders",
             obsessive_compulsive_disorders_numeric = "OCDs",
             psychotic_disorders_numeric = "Psychotic disorders",
             mhd.mania_hypomania_bipolar_or_manicdepression_numeric = "Bipolar disorders",
             mhd.posttraumatic_stress_disorder_ptsd_numeric = "PTSD",
             mhd.autism_aspergers_or_autistic_spectrum_disorder_numeric = "Autism",
             mhd.attention_deficit_hyperactivity_disorder_numeric = "ADHD",
             mhd.personality_disorder_numeric = "Personality disorders",
             control_numeric = "Controls"
           ))

head(dat.long.psych)
```

@@

```{r Disorder by Gender}
Disorder_by_Gender <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Gender) %>%
  count(Disorder)

Disorder_by_Gender
```

```{r Disorder by Gender plot}
Disorder_by_Gender %>%
  ggplot(mapping = aes(
    x = Disorder,
    y = n)
    ) +
  geom_bar(
    aes(fill = Gender),
#    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorder",
    y = "Number of participants",
    fill = "Gender") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  coord_flip()
```

```{r Disorder by Sample}
Disorder_by_Sample <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Sample) %>%
  count(Disorder)
```

```{r Disorder by Sample plot}
Disorder_by_Sample %>%
  ggplot(mapping = aes(
    x = Disorder,
    y = n)
    ) +
  geom_bar(
    aes(fill = Sample),
#    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorder",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette4,
#    labels = c("Prepandemic", "Retrospective")
    ) +
  coord_flip()
```

Ethnicity should be recoded
European
Mixed or multiple ethnic origins
African or African British
Asian or Asian British
Arab
```{r Disorder by Ethnicity}
Disorder_by_Ethnicity <- dat.long.psych %>%
  filter(Diagnosis == 1) %>%
  group_by(Disorder, Ethnicity) %>%
  count(Disorder)
```

```{r Disorder by Ethnicity plot}
Disorder_by_Ethnicity %>%
  ggplot(mapping = aes(
    x = Disorder,
    y = n)
  ) +
  geom_bar(
    aes(fill = Ethnicity),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Disorder",
    y = "Number of participants",
    fill = "Ethnicity") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6,
    #    labels = c("Prepandemic", "Retrospective")
  ) +
  coord_flip()
```

@@

# Ethnicity
```{r Ethnicity per Sample}
Ethnicity_by_Sample <- dat %>%
  group_by(Ethnicity, Sample) %>%
  count(Ethnicity)

Ethnicity_by_Sample
```

```{r Ethnicity by Sample plot}
Ethnicity_by_Sample %>%
  ggplot(mapping = aes(
    x = Ethnicity,
    y = n)
  ) +
  geom_bar(
    aes(fill = Sample),
    #    position = "dodge",
    stat="identity",
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Ethnicity",
    y = "Number of participants",
    fill = "Sample") +
  theme_personal +
  scale_fill_manual(
    values = COPINGpalette6,
    #    labels = c("Prepandemic", "Retrospective")
  ) +
  coord_flip()
```

# Gender
```{r Gender per Sample}
dat %>%
  group_by(Gender, Sample) %>%
  count(Gender)
```

# highest_education
```{r highest_education per Sample}
dat %>%
  group_by(highest_education, Sample) %>%
  count(highest_education)
```

# Key worker status
```{r Key worker status per sample}
dat %>%
  group_by(key_worker, Sample) %>%
  count(key_worker)
```

# COVID
## COVID swab test positive
```{r COVID positive swap}
dat %>%
  group_by(respiratory.throat_swab_test_nose, Sample) %>%
  count(respiratory.throat_swab_test_nose)
```

## antibody test
```{r COVID positive antibody test}
dat %>%
  group_by(respiratory.results_antibody_test, Sample) %>%
  count(respiratory.results_antibody_test)
```

## COVID probable diagnosis
```{r COVID probable diagnosis}
dat %>%
  group_by(covid_probable_case_incl_screening, Sample) %>%
  count(covid_probable_case_incl_screening)
```

# Questionnaire comparions

## gad
```{r gad gather to long format Time_points}
dat.long.gad.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "gad.diff_score",
    all_of(gad.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Time_point_raw,
    gad.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "gad.diff_score_", "")) %>%
  select(-Time_point_raw)

@@

makes this into percentages
```{r gad absolut plot}
dat.long.gad.diff %>%
  ggplot(mapping = aes(y = gad.diff_score)) +
  geom_bar() +
  theme_minimal() +
  facet_wrap(facets = vars(Time_point))
```

```{r gad percentage plot prepan_retro}
dat.long.gad.diff %>%
  filter(Time_point != "prepan_retro") %>%
  ggplot(mapping = aes(
    x = gad.diff_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Difference GAD7 score",
    y = "Percentage of participants",
    fill = "Comparison:\nBaseline vs") +
  theme_personal +
  scale_fill_manual(values = palette, labels = c("Prepandemic", "Retrospective"))
```

```{r gad plot prepan_retro diff}
dat.long.gad.diff %>%
  filter(Time_point == "prepan_retro") %>%
  ggplot(mapping = aes(
    x = gad.diff_score,
    y = stat(density*width),
    fill = Time_point)) +
  geom_histogram(
    binwidth = 1,
    position = position_dodge2(preserve = "single")
  ) +
  labs(
    x = "Difference GAD7 score",
    y = "Percentage of participants",
    fill = "Comparison:\nPrepandemic vs") +
  theme_personal +
  scale_fill_manual(values = palette, labels = c("Retrospective"))
```

# phq

```{r phq gather to long format Time_points}
dat.long.phq.diff <- dat %>%
  gather(
    key = "Time_point_raw",
    value = "phq.diff_score",
    all_of(phq.diff_score_cols)) %>%
  select(
    ID,
    Sample,
    Gender,
    Ethnicity,
    Time_point_raw,
    phq.diff_score
  ) %>%
  mutate(Time_point = str_replace(Time_point_raw, "phq.diff_score_", "")) %>%
  select(-Time_point_raw) %>%
  drop_na(phq.diff_score)
